### 2019-03-13 streaming_data_marketplace.tex
Add FTC template

\documentclass{svproc}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\def\UrlFont{\rmfamily}

\begin{document}
\mainmatter              % start of a contribution
%
\title{Data Marketplace for Streaming Data}
%
\titlerunning{Streaming Data Marketplace}  % abbreviated title (for running head)
%
\author{Ching-Chun Huang \and Jing-Hua Lin
}
%
\authorrunning{Ching-Chun Huang et al.} % abbreviated author list (for running head)
\tocauthor{Ching-Chun Huang, Jing-Hua Lin}
%
\institute{Cheng-Kung University, Tainan, Taiwan,\\
\email{jserv.tw@gmail.com}
}

\maketitle              % typeset the title of the contribution

%at least 70 words, at most 150 words
%It will be set in 9-point font size and be inset 1.0 cm from the right and left margins.
\begin{abstract}
abstract here
\keywords{streaming data, crowd sensing, data marketplace}
\end{abstract}

\section{Introduction}

\section{Related Work}

% ---- Bibliography ----
\begin{thebibliography}{1}
%
%\bibitem {smit:wat}
%Smith, T.F., Waterman, M.S.: Identification of common molecular subsequences.
%J. Mol. Biol. 147, 195?197 (1981). \url{doi:10.1016/0022-2836(81)90087-5}

\end{thebibliography}
\end{document}